### 2020-12-10 Thesis-Chapter2-Background.tex
Editorial improvements

\Ac{cryptography} is the study of techniques for communicating secretly in the presence of third parties. This is
performed by using a \ac{cipher} to translate between \ac{plaintext} and \ac{ciphertext}. A cipher is a tool or
algorithm that performs the translation, plaintext is the original data, and ciphertext is the encoded data. The
processes for translating from plaintext to ciphertext and back are \ac{encryption} and \ac{decryption}. A well-designed
cipher ensures that only those parties with the correct secret information, the \ac{key}, can perform encryption or
decryption on the text.

@@

There are two major cryptographic protocol families, symmetric and asymmetric. In \ac{sym-crypto}, which was invented
long ago, encryption and decryption are performed with the same key, and both parties must have this key in order to
communicate securely. In \ac{asym-crypto}, which was invented recently and is also called \ac{public-key-crypto},
encryption and decryption are performed with two paired keys, called the public key and the private key. The innovation
of \ac{public-key-crypto} is that the public key is not secret, and using this technique, two parties can communicate
securely without requiring a previously agreed upon shared secret.

One can also define two major cryptographic applications---securing \acf{DIM} and \acf{DAR}. Each application is a
different problem that results in different solutions. \Ac{DIM} typically uses long-lived \ac{asym-crypto} keys to
perform authentication and establish short-lived \ac{sym-crypto} session keys; the session keys perform the actual
encryption of the data in motion. Two important properties of \ac{DIM} encryption protocols are \ac{forward-secrecy} and
\ac{replay-protection}. \Ac{forward-secrecy} ensures that a leaked private key or session key does not compromise any
other private key or session key; \ac{replay-protection} ensures that messages cannot be replayed by an attacker without
detection \cite{bellovin_thinking_2016}. \Ac{E2EE} for instant messaging services is an example of encryption for
\ac{DIM}.

\Ac{DAR} by nature must use long-lived keys for encryption. Instead of being negotiated and randomly generated at
encryption-time, as \ac{DIM} session keys are, these keys are either derived from a user-entered password or are stored
somewhere in computer memory, which often takes place with the assistance of dedicated hardware. \Ac{disk-encryption}
for laptops and mobile devices is an example of encryption for \ac{DAR}.

Data secrecy is not cryptography's only use, however. Precisely speaking, data secrecy is the purpose of encryption.
Encryption is cryptography's ``killer app,'' but it is only a part of cryptography's usefulness to security.

Cryptography is the technical foundation for many forms of computer and network security. Security is defined in terms
of several properties: authentication, integrity, non-repudiation, confidentiality, availability, and authorization
\cite{shostack_threat_2014}. These system-level properties emerge from both the composition of the system's components
(architecture) and the security of the components themselves; likewise, the security of each component emerges from the
architecture of its subcomponents and the security of the subcomponents themselves. At the end of this chain of
analysis, the security of primitive components commonly relies on cryptography.

\mytab{table-security-and-crypto} connects each security property to its cryptographic basis. While lacking direct
cryptographic foundations, availability indirectly relies on each other property, and authorization schemes are
typically rooted in authentication.

@@

Violations of cryptographic integrity in primitive components, introduced by design or by accident, could have
catastrophic effects. Compromised digital certificates, hashes, or encryption would enable spoofing, tampering, and
information disclosure done both for their own sake and as tools in general attacks. The importance of taking these
tools away is clear when such attacks include interference with elections \cite{mueller_2018}, multi-billion dollar
disruptions of business \cite{greenberg_2018_notpetya}, and hospital network ransoms that disrupt care of patients
\cite{goodin_ransomware_2019}.

@@

Encryption's dual role as enabler of privacy and cornerstone of security is at the heart of the EA debate. Cryptography
has an absolute, mathematical power---a power that is objectionable in the realm of privacy, but necessary in the realm
of security.

@@

Computerized encryption is a new technology, but the field of cryptography is old, since demand for privacy is as old as
communication itself. The most well known ancient example of rudimentary encryption is the Caesar cipher, named after
the character substitution technique Julius Caesar used to protect private correspondence
\cite{luciano_cryptology_1987}. Significant developments include the first formal cryptographic study by Arab scholars
in the eighth century, advancements made out of necessity in the twentieth century's world wars, and the application of
computers to cryptographic problems \cite{kahn_codebreakers_1996}. Claude Shannon's formalized the modern ``mathematical
analysis of cryptography'' in 1949 \cite{shannon_communication_1949}, and Whitfield Diffie and Martin Hellman published
research on \ac{public-key-crypto} in 1976 \cite{diffie_new_1976}. The discovery of \ac{public-key-crypto} was an
important advance. Combined with computer networking and personal computing advances in the 1980s, it for the first time
put the power of strong encryption not just in the hands of governments and militaries, but of ordinary people.

However, this power enables very strong privacy, and privacy is a property governments aim to regulate. The use of
encryption for for private purposes has a contentious history. In 1587, Mary Queen of Scots was convicted of treason
based on evidence from decrypted letters, and in 1807, prosecutors trying Aaron Burr for treason tried to force
testimony from his secretary on the contents of encrypted messages \cite{kerr_encryption_2017}. Government's natural
stance towards privacy makes it inherently uncomfortable with the rapid increase in availability of strong encryption.

% Electronic encryption must be understood as a technological development embedded in the longer history of privacy and
% human rights.

@@

All these realities---government discomfort with absolute privacy, rapidly increasing availability of strong encryption,
and a blossoming computer industry foundationally reliant on cryptography---came to a head in what has become known as
the first ``crypto war.''

In 1976, the same year Diffie and Hellmann published their research in public key cryptography, the U.S. Congress passed
the \ac{AECA} and declared that strong cryptography is subject to export controls \cite{kehl_right_2015}. The opening
shots of \ac{the-cw1} came in 1991, when the U.S. Senate introduced, but did not pass, a bill mandating access to plain
text contents when authorized by law. In response, Philip Zimmermann released \ac{PGP}, a public key encryption tool to
secure email, in order that strong cryptography be ``made available to the American public before it became illegal to
use'' \cite{zimmermann_1996}. In 1993, the Clinton administration introduced the \ac{clipperchip} \cite{press_1993} with
the goal of ``providing the public with strong cryptographic tools without sacrificing the ability of law enforcement
and intelligence agencies to access unencrypted versions of those communications'' \cite{thompson_2015}. Citing the
foundational role of cryptography in security---and the potential of human rights abuses for compromised
privacy---industry and prominent technical leaders reacted negatively to the initiative \cite{kehl_right_2015}
\cite{zimmermann_1996}. When a prominent security researcher discovered flaws that allowed users to subvert the
\ac{clipperchip} mechanisms \cite{blaze_protocol_1994}, the proposal died.

Despite the failure of the \ac{clipperchip}, the debate over export controls and access to strong encryption continued,
primarily focused on various key escrow proposals \cite{thompson_2015}. In 1996, two pro-encryption bills were
introduced in the US Congress. In the Senate, it was S.1726, the \ac{PROCODEAct} of 1996, to abolish the export controls
on encryption software \cite{burns_s1726_1996}. In the House, it was HR.3011, the \ac{SAFEAct} of 1996, to explicitly
allow arbitrarily strong encryption for all legal activity \cite{goodlatte_hr3011_1996}. Indeed, the \ac{SAFEAct} went
further, including ``provisions that would have barred the government from creating a mandatory key escrow system and
would also have removed export restrictions on most generally available software and hardware containing encryption''
\cite{kehl_right_2015}. Zimmermann, the author of PGP, who had by this time endured an investigation by the Customs
Service for publishing his work, testified before the Senate in favor of the \ac{PROCODEAct} \cite{zimmermann_1996}. In
November of that year, the Clinton administration released an executive order effectively removing export controls on
strong encryption products along the lines of the \ac{PROCODEAct}, which never came to pass \cite{clinton_1996}.

From 1996 to 1999, the \ac{SAFEAct} was proposed several times, had hearings, and gained support \cite{kehl_right_2015}.
By this time, there was ``an overwhelming amount of evidence against moving ahead with any key escrow schemes''
\cite{thompson_2015}, and in 1999 the Clinton administration abruptly changed course, adopting almost all \ac{SAFEAct}
proposals \cite{kehl_right_2015}. This development marks the end of \ac{the-cw1}.

@@

Two developments paved the way for \ac{the-cw2}. The first was actually a non-development: when the Clinton
administration changed its encryption policy, the U.S. House dropped the \ac{SAFEAct} \cite{goodlatte_hr3011_1996}. The
\ac{PROCODEAct} and other encryption-related bills went unpassed as well, as elaborated in \mysec{sec-us-reg-fails}.
This legislative failure meant that rather than being written in the pen of law, encryption policy was written in the
pencil of executive order.

The second development was the events of and reaction to the morning of September 11, 2001. These events occurred in the
context of decades of all three branches of government expanding law enforcement powers. Presidents promoted the idea of
being ``tough on crime'' and launched a war on drugs; legislators passed bills such as 1978's \ac{FISA}, which provided
broad surveillance powers over foreign nationals, and in some cases U.S. citizens \cite{rodino_1978}; and the judiciary
practiced a ``jurisprudence of crime control'' that gave police broad leeway \cite{gizzi_fourth_2016}.

In this context of expanded surveillance, the scale and audacity of the September 11 attacks seemed to justify yet
greater investigatory powers \cite{bloss_escalating_2007}. October of 2001 saw passage of the USA PATRIOT Act, which had
several significant impacts: it weakened the limitation for using \ac{FISA} requests on U.S. citizens; it expanded the
scope of what could be compelled via \ac{FISA} orders; it authorized ``roving'' wiretaps; and it increased the power of
\acp{NSL}, which can be used without judicial review to compel information from digital service providers while
precluding any public disclosure of the event \cite{sensenbrenner_2001} \cite{shamsi_2011}. The government was
performing more surveillance than ever, and even these weakened limitations would be repeatedly violated
\cite{shamsi_2011} \cite{tucker_2020}.

The weakness of data protection laws and absence of encryption protection laws left the door open for a second crypto
war, and increased government surveillance set events in motion toward conflict, but it was the 2013 Snowden revelations
that constituted the crossing of the threshold. Though policies such as the U.S. \ac{NSA}'s warrantless wiretapping
caused a stir, it was the mass data collection under the agency's \ac{PRISM} and related programs that caused the real
public outcry \cite{landau_making_2013}. With the public interest focused on digital privacy, encryption promised a
technical solution. Benefiting from advancements in underlying technologies and under pressure from both the human
rights field and from ordinary users demanding improved privacy, U.S. tech companies responded by introducing default
device encryption for \acl{DAR} and \acl{E2EE} for \acl{DIM} \cite{treguer_us_2018}.

@@

The U.S. corporate response to Snowden should not be overly construed as a morally motivated defense of civil rights, as
their behavior is a matter of several practical factors \cite{treguer_us_2018}, not least of which is the pursuit of
basic digital security. Recall encryption and cryptography's critical role in security as described in
\mysec{sec-crypto-basics}, and that encryption's dual role as enabler of strong privacy and cornerstone of computer
security is a central conflict in the \ac{EA} debate. When questioned by lawmakers, technical leaders and executives
have repeatedly appealed to the fact that strong cryptography is a necessity for security \cite{schulze_clipper_2017}.

@@

\paragraph*{``Security vs. Security''} The relationship between security and privacy is sometimes viewed as
antagonistic, an assumption implicit in the ``nothing-to-hide'' argument, but discussions in recent years have shown
that the debate is more about ``Security vs. Security'' than ``Security vs. Privacy''
\cite{stalla_bourdillon_privacy_2014}. In this context, there are two types of security: public security, the pursuit of
law enforcement, and cybersecurity, the pursuit of the technical community. Both sides of the debate have by now
acknowledged that privacy and public security are not always in conflict, and in fact \ac{EA}'s effects on public
security vs. cybersecurity---which in return affects public security---may be more important \cite{schneier_2019}.

\paragraph*{Apple vs. FBI} The 2015 San Bernardino terror attack in which the suspect was killed, leaving behind a
locked iPhone, resulted in the first major battle of \ac{the-cw2}. The \ac{FBI} issued an order under the \acrlong{AWA}
to compel Apple to unlock the device, which was among the first generation of Apple's fully encrypted iPhones; Apple
objected on grounds that it was ``unreasonably burdensome'' and would undercut the integrity of all iOS devices
\cite{schulze_clipper_2017}. The case occupied U.S. District courts, the media, and the political world's attention from
February 16, 2016, when the warrant was issued, to March 28, when the \ac{FBI} announced they had gained access to the
phone through alternate means \cite{novet_2016}. Even though the phone proved not to have any important data
\cite{schulze_clipper_2017}, the high profile case featured two characteristic elements of the new crypto
wars---terrorism and device encryption. Shortly after the case ended, a pair of Republican and Democratic senators
jointly released a discussion draft of a bill that would have forced Apple to comply \cite{burr_2016}. The draft did not
make it to the floor, and the unresolved nature of the court case left the legal debate unresolved.

\paragraph*{The \Ac{horsemen}} In the 1990s, an influential encryption advocate coined the phrase ``\ac{horsemen-full}'' to
describe four reasons that governments and law enforcement agencies use to undercut public support for strong encryption
\cite{may_1994}. The traditional list includes terrorists, pedophiles, drug dealers, and money launderers, though
sometimes kidnappers are substituted for money launderers \cite{schneier_scaring_2019}. Reflecting the times,
\ac{the-cw1} emphasized drug trafficking whereas \ac{the-cw2} is emphasizing terrorism and child pornography (\ac{CSAM})
\cite{schulze_clipper_2017}. The focus on terrorism is likely a result of culture that emerged from the September 11
terror attacks as well as the San Bernardino terror attack that launched the Apple vs. \ac{FBI} case; the focus on
\ac{CSAM} is likely a result of the growth of the ``dark web'' and a revelatory 2019 investigation by the \textit{New
York Times} \cite{keller_internet_2019}.

\paragraph*{\ac{DIM} vs. \ac{DAR}} As noted in \mysec{sec-crypto-basics}, securing \acl{DIM} and \acl{DAR} are two
different problems that result in different solutions. Therefore, both technical and policy proposals often split their
recommendations along these lines \cite{group_2019} \cite{owen_law_2018}. The focus from law enforcement in the U.S. has
pendulated between each of these applications a few times \cite{schneier_2019}. The typical \ac{DIM} application is
\ac{E2EE} in instant messaging applications, whereas the typical \ac{DAR} application is \ac{disk-encryption} of mobile
phones. A research group at Carnegie Mellon with participants from both sides of the debate has identified \ac{EA} for
\ac{DAR} as the more tractable problem \cite{group_2019}.

\paragraph*{Volatile Politics} Finally, \ac{the-cw2} is taking place amongst a political landscape marked by
uncertainty, extremism, misinformation, and a pandemic that will have unknown long-term impacts. Since legislature is
part of the broader data security threat model, its behavior is important. Unfortunately, the present political dynamics
and the lack of reliable data on what law enforcement agencies need make it difficult for the technical community to
engage productively with policymakers \cite{granick_2018}.

That brings us to where the debate stands today. A December 2019 U.S. Senate Judiciary Committee featured the top two
U.S. parties both expressing anger over whether data should be ``beyond the reach of the law'' and expressed threats
such as ``get your act together, or we will gladly get your act together for you'' \cite{geller_2019}. In March 2020,
the Republican atop the Senate Judiciary Committee introduced S.3398, the \ac{EARNITAct} of 2020, which seeks to
establish a 16-chair committee with the power to revoke online platforms' liability protection for user-submitted
content if the committee determines the platform is not doing enough to fight \ac{CSAM}, a legal synonym for child
pornography \cite{graham_s3398_2020}. As illustrated by the previously mentioned \textit{New York Times} feature
\cite{keller_internet_2019} that likely triggered the December 2019 hearing, the problem is enormous; however, critics
largely saw the \ac{EARNITAct} as an attempt to indirectly yet effectively outlaw \ac{E2EE} \cite{newman_2020}
\cite{pfefferkorn_2020}. In response, legislators narrowed the focus of the \ac{EARNITAct}, but only after unveiling the
\ac{LAEDAct}

@@

There are several technical approaches to \ac{EA}. This section categorizes these approaches according to one of three
basic approaches---\ac{key-escrow}, cryptographic recovery without a key, or non-cryptographic means of acquiring
plaintext.

@@

The most obvious approach to \ac{EA} is to store additional copies of the encrypting keys. A variation is to store
information that can be used to derive additional copies. This approach is called \ac{key-escrow}. There are several
classes of key escrow.

@@

Some schemes compensate for the risk of concentrated sensitive information by massively distributing the secret
information. In this arrangement, trust is put in the distributed system, not in any single party. Such systems offer
high data availability while making it difficult for attempts to acquire keys to be secret or to subvert access
policies. This approach was introduced in \ac{the-cw1} \cite{goos_oblivious_1996} and has modern blockchain
\cite{phan_key_2017} and distributed device \cite{servan_schreiber_jje_2020} variations. These approaches are not as
thoroughly explored as trusted-party \ac{key-escrow}.

@@

A third type of \ac{key-escrow} is device-based. In device-based \ac{key-escrow}, special key information is held not by
trusted parties or distributed systems, but on the device itself in specialized hardware. Advances in secure hardware
``enclaves'' since \ac{the-cw1} have enabled secure local key storage. Approaches relying on hardware primarily target
\ac{DAR}, and may include time elements and split keys with trusted parties as well \cite{savage_lawful_2018}
\cite{ozzie_2018}.

@@

Research in \ac{the-cw1} included the introduction of ``translucent'' cryptography, in which an observer could surveil
some but not all communications using statistical cryptographic methods \cite{bellare_translucent_1996}. Under this
scheme, no keys are escrowed, but law enforcement could recover a predetermined fraction---say, 40\%---of plaintext
data. That fraction would be designed to balance privacy, digital security, and public security. This research, though
decidedly more experimental than key escrow proposals, was explicitly pursued to demonstrate that alternatives to key
escrow do exist \cite{bellare_translucent_1996}.

@@

A recent proposal adopts the idea of crumple zones from automotive engineering---``in an emergency situation the
construction should break a little bit in order to protect the integrity of the system as a whole and the safety of its
human users'' \cite{wright_crypto_2018}. Similar to the translucent approach, no keys are escrowed and only passive
surveillance is possible. Strong cryptography is a precondition for this scheme, in which keys are generated in a manner
that makes them inherently recoverable only after a large up front and marginal cost.

>>>>

\Ac{cryptography} is the study of techniques used to communicate securely in the presence of third parties. This is
performed by using a \ac{cipher} to translate between \ac{plaintext} and \ac{ciphertext}. A cipher is a tool or
algorithm that performs the translation, plaintext is the original data, and ciphertext is the encoded data.
\Ac{encryption} and \ac{decryption} are the processes for translating from plaintext to ciphertext and back. A
well-designed cipher ensures that only those parties with the correct \ac{key}, or secret information, can perform
encryption or decryption on the text.

@@

There are two major cryptographic protocol families, symmetric and asymmetric. Symmetric protocols have been around for
thousands of years, but asymmetric protocols were only invented in recent decades. In \ac{sym-crypto}, encryption and
decryption are performed with the same key, and both parties must have this key in order to communicate securely. In
\ac{asym-crypto}, also known as \ac{public-key-crypto}, encryption and decryption are performed with two paired keys,
called the public key and the private key. The public key is not secret, but due to its relationship with the private
key, it can be used to establish identity and encrypted communication. Using this technique, two parties can communicate
privately without requiring previously agreed-upon secret information.

There are also two major cryptographic applications, securing \ac{DIM} and \ac{DAR}. Each application presents different
challenges that require different solutions. \Ac{DIM} typically uses long-lived \ac{asym-crypto} keys to perform
authentication and establish short-lived \ac{sym-crypto} session keys. The session keys perform the actual encryption of
the \acl{DIM}. \Ac{forward-secrecy} and \ac{replay-protection} are two important properties of \ac{DIM} encryption
protocols. \Ac{forward-secrecy} ensures that a leaked private key or session key does not compromise any other private
key or session key; \ac{replay-protection} ensures that messages cannot be replayed by an attacker without detection
\cite{bellovin_thinking_2016}. \Ac{E2EE} for instant messaging services is an example of encryption for \ac{DIM}.

\Ac{DAR} by nature must use long-lived keys for encryption. Instead of being negotiated and randomly generated at
encryption-time, as \ac{DIM} session keys are, these keys are either derived from a user-entered password or are stored
somewhere in computer memory. Secure storage often takes place with the assistance of dedicated hardware.
\Ac{disk-encryption} for laptops and mobile devices is an example of encryption for \ac{DAR}.

Data secrecy is not cryptography's only use, however. Data secrecy is the purpose of encryption, but encryption is only
one application of cryptography. Although encryption is cryptography's ``killer app,'' it is just a part of
cryptography's usefulness to security.

Cryptography is the technical foundation for many forms of computer and network security. Security is defined in terms
of several properties: authentication, integrity, non-repudiation, confidentiality, availability, and authorization
\cite{shostack_threat_2014}. These system-level properties emerge from both the composition of the system's components
(architecture) and the security of the components themselves. Likewise, the security of each component emerges from the
architecture of its subcomponents and the security of the subcomponents themselves. At the end of this chain of
analysis, the security of primitive components commonly relies on cryptography.

\mytab{table-security-and-crypto} connects each security property to its cryptographic basis. While lacking direct
cryptographic foundations, the property of availability indirectly relies on the other properties, and authorization
schemes are typically rooted in authentication.

@@

Violations of cryptographic integrity in primitive components, introduced by design or by accident, could have
catastrophic effects. Compromised digital certificates, hashes, or encryption would enable spoofing, tampering, and
information disclosure done both for their own sake and as tools in general attacks. The importance of neutralizing
these tools is clear when such attacks include interference with elections \cite{mueller_2018}, multi-billion dollar
disruptions of business \cite{greenberg_2018_notpetya}, and hospital network ransoms that disrupt care of patients
\cite{goodin_ransomware_2019}.

@@

Encryption's dual role as an enabler of privacy and cornerstone of security is at the heart of the EA debate.
Cryptography has an absolute, mathematical power---a power that is objectionable in the realm of privacy, but necessary
in the realm of security.

@@

Computerized encryption is a new technology, but the field of cryptography is old, since demand for privacy is as old as
communication itself. The most well-known ancient example of rudimentary encryption is the Caesar cipher, named after
the character substitution technique Julius Caesar used to protect private correspondence
\cite{luciano_cryptology_1987}. Significant developments include the first formal cryptographic study by Arab scholars
in the eighth century, advancements made out of necessity in the twentieth century's world wars, and the application of
computers to cryptographic problems \cite{kahn_codebreakers_1996}. Claude Shannon formalized the modern ``mathematical
analysis of cryptography'' in 1949 \cite{shannon_communication_1949} and Whitfield Diffie and Martin Hellman published
research on \ac{public-key-crypto} in 1976 \cite{diffie_new_1976}. The discovery of \ac{public-key-crypto} was an
important advance. Combined with computer networking and personal computing advances in the 1980s, it put the power of
strong encryption not just in the hands of governments and militaries, but of ordinary people.

However, this power enables very strong individual privacy. Government is inherently uncomfortable with individual
privacy; therefore, the use of encryption for private purposes has a contentious history. In 1587, Mary Queen of Scots
was convicted of treason based on evidence from decrypted letters, and in 1807, the prosecutors who tried Aaron Burr for
treason attempted to force testimony from his secretary on the contents of encrypted messages
\cite{kerr_encryption_2017}.

@@

These circumstances---government discomfort with absolute privacy, rapidly increasing availability of strong encryption,
and a blossoming computer industry foundationally reliant on cryptography---came to a head in what has become known as
the first ``crypto war.''

In 1976, Diffie and Hellmann published their seminal research in public key cryptography. That same year, the U.S.
Congress passed the \ac{AECA} and declared that strong cryptography is subject to export controls
\cite{kehl_right_2015}. \Ac{the-cw1} began in 1991, when the U.S. Senate introduced, but did not pass, a bill mandating
access to plain text contents when authorized by law. In response, Philip Zimmermann released \ac{PGP}, e-mail
encryption software, in order that strong cryptography be ``made available to the American public before it became
illegal to use'' \cite{zimmermann_1996}. In 1993, the Clinton administration introduced the \ac{clipperchip}
\cite{press_1993} with the goal of ``providing the public with strong cryptographic tools without sacrificing the
ability of law enforcement and intelligence agencies to access unencrypted versions of those communications''
\cite{thompson_2015}. Citing the foundational role of cryptography in security---and the potential of human rights
abuses for compromised privacy---industrial and technical leaders reacted negatively to the initiative
\cite{kehl_right_2015} \cite{zimmermann_1996}. When a prominent security researcher discovered flaws that allowed users
to subvert the \ac{clipperchip} mechanisms \cite{blaze_protocol_1994}, the proposal died.

Despite the failure of the \ac{clipperchip}, the debate over export controls and access to strong encryption continued,
primarily focused on various key escrow proposals \cite{thompson_2015}. In 1996, two pro-encryption bills were
introduced in the U.S. Congress, one in the Senate and one in the House. The Senate saw S.1726, the \ac{PROCODEAct} of
1996, which sought to abolish the export controls on encryption software \cite{burns_s1726_1996}. The House saw HR.3011,
the \ac{SAFEAct} of 1996, which similarly sought to remove export controls, but further sought to explicitly allow
arbitrarily strong encryption for all legal activity and preclude mandatory \ac{EA} schemes such as the \ac{clipperchip}
\cite{goodlatte_hr3011_1996}. Zimmermann, the author of PGP, who had by this time endured an investigation by the
Customs Service for publishing his work, testified before the Senate in favor of the \ac{PROCODEAct}
\cite{zimmermann_1996}. In November of 1996, the Clinton administration released an executive order effectively removing
export controls on encryption products along the lines of the \ac{PROCODEAct}, which never came to pass
\cite{clinton_1996}.

From 1996 to 1999, the \ac{SAFEAct} was proposed several times, was discussed at hearings, and gained support
\cite{kehl_right_2015}. By this time, there was ``an overwhelming amount of evidence against moving ahead with any key
escrow schemes'' \cite{thompson_2015}, and in 1999 the Clinton administration abruptly changed course, adopting almost
all \ac{SAFEAct} proposals \cite{kehl_right_2015}. This development marked the end of \ac{the-cw1}.

@@

Two developments paved the way for \ac{the-cw2}. The first was the result of inaction: when the Clinton administration
changed its encryption policy, the U.S. House dropped the \ac{SAFEAct} \cite{goodlatte_hr3011_1996}. The \ac{PROCODEAct}
and other encryption-related bills went unpassed as well, as elaborated in \mysec{sec-us-reg-fails}. This legislative
failure meant that rather than being written in the pen of law, encryption policy was written in the pencil of executive
order.

The second development was the terror attacks of September 11, 2001. The events of that day occurred in the context of
expanding law enforcement powers fueled for decades by all three branches of government. Presidents promoted the idea of
being ``tough on crime'' and launched a war on drugs; legislators passed bills such as 1978's \ac{FISA}, which provided
broad surveillance powers over foreign nationals, and in some cases U.S. citizens \cite{rodino_1978}; and the judiciary
practiced a ``jurisprudence of crime control'' that gave police broad leeway \cite{gizzi_fourth_2016}.

In this context of expanded surveillance, the scale of the September 11 attacks seemed to justify yet greater
investigatory powers \cite{bloss_escalating_2007}. October of 2001 saw passage of the USA PATRIOT Act, which had several
significant impacts: (1) it weakened the limitation for using \ac{FISA} requests on U.S. citizens; (2) it expanded the
scope of what could be compelled via \ac{FISA} orders; (3) it authorized ``roving'' wiretaps; and (4) it increased the
power of \acp{NSL}, which can be used without judicial review to compel information from digital service providers while
precluding any public disclosure of the event \cite{sensenbrenner_2001} \cite{shamsi_2011}. The government was
performing more surveillance than ever, and even these weakened limitations would be repeatedly violated
\cite{shamsi_2011} \cite{tucker_2020}.

The weakness of data protection laws and absence of encryption protection laws left the door open for a second crypto
war. Increased government surveillance set events in motion toward conflict. These two developments prepared the way,
but it was the 2013 Snowden revelations that constituted the crossing of the threshold. Though policies such as the U.S.
\ac{NSA}'s warrantless wiretapping caused a stir, it was the mass data collection under the agency's \ac{PRISM} and
related programs that caused the real public outcry \cite{landau_making_2013}. With the public interest focused on
digital privacy, encryption promised to be a technical solution. U.S. tech companies responded by introducing default
device encryption for \acl{DAR} and \acl{E2EE} for \acl{DIM}.

@@

The U.S. corporate response to Snowden should not be overly construed as a morally motivated defense of civil rights.
Their behavior is a matter of several practical factors, including market pressure, cultural pressure, and the
advancement in available encryption technology \cite{treguer_us_2018}. Another factor is the pursuit of basic digital
security. Recall encryption and cryptography's critical role in security as described in \mysec{sec-crypto-basics}. When
lawmakers question technical leaders about their companies' encryption policies, executives have repeatedly appealed to
the fact that strong cryptography is a necessity for security \cite{schulze_clipper_2017}.

@@

\paragraph*{``Security vs. Security''} The relationship between security and privacy is sometimes viewed as
antagonistic, an assumption implicit in the ``nothing-to-hide'' argument. However, discussions in recent years have
shown that the debate is more about ``Security vs. Security'' than ``Security vs. Privacy''
\cite{stalla_bourdillon_privacy_2014}. In this context, there are two types of security. First, public security (often
referred to as ``safety'' in this paper), which is the pursuit of law enforcement. Second, cybersecurity, which the
pursuit of the technical community. Both sides of the debate have by now acknowledged that privacy and public security
are not always in conflict. In fact \ac{EA}'s effects on public security vs. cybersecurity---which in return affects
public security---may be more important \cite{schneier_2019}.

\paragraph*{Apple vs. FBI} The 2015 San Bernardino terror attack resulted in the first major battle of \ac{the-cw2}. The
suspect was killed, leaving behind a locked iPhone. The \ac{FBI} issued an order under the \acrlong{AWA} to compel Apple
to unlock the device, which was among the first generation of Apple's fully encrypted iPhones. Apple objected on grounds
that it was ``unreasonably burdensome'' and would undercut the integrity of all iOS devices \cite{schulze_clipper_2017}.
The case occupied U.S. District courts, the media, and the attention of the political world from February 16, 2016, when
the warrant was issued, to March 28, when the \ac{FBI} announced they had gained access to the phone through alternate
means \cite{novet_2016}. Although the phone proved not to contain any important data \cite{schulze_clipper_2017}, the
high profile case featured two characteristic elements of the new crypto wars---terrorism and device encryption. Shortly
after the case ended, a pair of Republican and Democratic senators jointly released a discussion draft of a bill that
would have forced Apple to comply \cite{burr_2016}. The draft did not make it to the floor, and the unresolved nature of
the court case left the legal debate unresolved.

\paragraph*{The \Ac{horsemen}} In the 1990s, an influential encryption advocate coined the phrase ``\ac{horsemen-full}''
to describe four reasons that governments and law enforcement agencies use to undercut public support for strong
encryption \cite{may_1994}. The original list includes terrorists, pedophiles, drug dealers, and money launderers,
though many substitute kidnappers for money launderers \cite{schneier_scaring_2019}. Reflecting the times, \ac{the-cw1}
emphasized drug trafficking whereas \ac{the-cw2} emphasizes terrorism and child pornography (increasingly referred to as
\ac{CSAM}) \cite{schulze_clipper_2017}. The focus on terrorism is likely a result of the September 11 terror attacks as
well as the San Bernardino terror attack that launched the Apple vs. \ac{FBI} case. The focus on \ac{CSAM} is likely a
result of the growth of the ``dark web'' and a revelatory 2019 investigation by the \textit{New York Times}
\cite{keller_internet_2019}.

\paragraph*{\ac{DIM} vs. \ac{DAR}} As noted in \mysec{sec-crypto-basics}, securing \acl{DIM} and \acl{DAR} are two
different challenges that require different solutions. Therefore, both technical and policy proposals often split their
recommendations along these lines \cite{group_2019} \cite{owen_law_2018}. Discussion of \ac{EA} for \ac{DIM} typically
involves \ac{E2EE}, whereas for \ac{DAR} it typically involves \ac{disk-encryption} of mobile phones. The focus of law
enforcement in the U.S. has pendulated between these several times \cite{schneier_2019}. A research group at Carnegie
Mellon with participants from both sides of the debate has identified \ac{EA} for \ac{DAR} as the more tractable problem
\cite{group_2019}.

\paragraph*{Volatile Politics} Finally, \ac{the-cw2} is taking place amongst a political landscape marked by
uncertainty, extremism, misinformation, and a pandemic that will have unknown long-term impacts. Since government is
part of the broader data security threat model, its behavior is important. Unfortunately, it is difficult for the
technical community to engage productively with policymakers due to present political dynamics and the lack of reliable
data showing what law enforcement agencies need \cite{granick_2018}.

That brings us to where the debate stands today. A December 2019 U.S. Senate Judiciary Committee featured the top two
U.S. parties both expressing anger over whether data should be ``beyond the reach of the law.'' Senators threatened tech
executives with statements such as ``get your act together, or we will gladly get your act together for you''
\cite{geller_2019}. In March 2020, the Republican atop the Senate Judiciary Committee introduced S.3398, the
\ac{EARNITAct} of 2020, which seeks to establish a 16-chair committee with the power to revoke online platforms'
liability protection for user-submitted content if the committee determines the platform is not doing enough to fight
\ac{CSAM} \cite{graham_s3398_2020}. As illustrated by the previously mentioned \textit{New York Times} feature
\cite{keller_internet_2019} that likely triggered the December 2019 hearing, the problem is enormous. However, critics
largely saw the \ac{EARNITAct} as an attempt to indirectly yet effectively outlaw \ac{E2EE} \cite{newman_2020}
\cite{pfefferkorn_2020}. In response, legislators narrowed the focus of the \ac{EARNITAct}, but only after unveiling the
\ac{LAEDAct}

@@

There are several technical approaches to \ac{EA}. This section organizes approaches into three
categories---\ac{key-escrow}, cryptographic recovery without a key, or non-cryptographic means of acquiring plaintext.

@@

Key escrow is the most obvious way to implement \ac{EA}. It involves storing additional copies of the encryption keys.
Alternatively, rather than directly storing copies, key escrow may involve storing information that can be used to
derive additional copies. Key escrow can be subdivided into various types according to where the escrowed information is
stored.

@@

Some schemes compensate for the risk of concentrated sensitive information by massively distributing the secret
information. In this arrangement, trust is put in the distributed system, not in any single party. Such systems offer
high data availability while making covert key recovery difficult. This approach was introduced in \ac{the-cw1}
\cite{goos_oblivious_1996} and has modern blockchain \cite{phan_key_2017} and distributed device
\cite{servan_schreiber_jje_2020} variations. These approaches are not as thoroughly explored as trusted-party
\ac{key-escrow}.

@@

The third type of \ac{key-escrow} is device-based. In device-based \ac{key-escrow}, special key information is held not
by trusted parties or distributed systems, but on the device itself in specialized hardware. Advances in secure hardware
``enclaves'' since \ac{the-cw1} have enabled secure local key storage. Approaches relying on hardware primarily target
\ac{DAR} and may include time elements and split keys with trusted parties as well \cite{savage_lawful_2018}
\cite{ozzie_2018}.

@@

Research in \ac{the-cw1} included the introduction of ``translucent'' cryptography, in which an observer could surveil
some but not all communications using statistical cryptographic methods \cite{bellare_translucent_1996}. Under this
scheme, no keys are escrowed, but law enforcement could recover a predetermined percentage of plaintext data. That
percentage would be designed to balance privacy, security, and safety. This research, though more experimental than key
escrow proposals, was explicitly pursued to demonstrate that alternatives to key escrow do exist
\cite{bellare_translucent_1996}.

@@

A recent proposal adopts the idea of crumple zones from automotive engineering---``in an emergency situation the
construction should break a little bit in order to protect the integrity of the system as a whole and the safety of its
human users'' \cite{wright_crypto_2018}. Similar to the translucent approach, no keys are escrowed and only passive
surveillance is possible. Keys are generated in a manner that makes messages inherently recoverable, but only after
extreme up front costs and significant marginal costs.
